\documentclass[12pt]{article}
\usepackage{braket}
\usepackage{amsmath}
\usepackage{listings}
\begin{document}
\title{Project 2}
\maketitle
\section{Method}



It can be shown that a unitary transformation of a vector conserves the dot product. 
The dot product between two of the basis vectors can be written as:

\begin{equation*}
\mathbf{v}_i^T \mathbf{v}_i=  \delta_{ij}
\end{equation*}
New vectors $\mathbf{w}_i$ are created by a unitary transformation of the basis vectors $\mathbf{v}_i$ 
\begin{equation*}
\mathbf{w}_i = \mathbf{Uv}_i
\end{equation*}
The dot product between the new vectors will be:
\begin{equation*}
\mathbf{w}_i^T\mathbf{w}_j  = (\mathbf{U}\mathbf{v}_i)^T\mathbf{U}\mathbf{v}_j = \mathbf{v}_i^T\mathbf{U}^T\mathbf{U}\mathbf{v}_j
\end{equation*}
Since $\mathbf{U}$ is a unitary matrix, 
\begin{equation*}
\mathbf{U}^T\mathbf{U} = \mathbf{1}
\end{equation*}
And so:
\begin{equation*}
\mathbf{w}_i^T\mathbf{w}_j  = \mathbf{v}_i^T \mathbf{v}_i=  \delta_{ij}
\end{equation*}
Thus, it has been proven that the dot product, and so also the orthogonality is conserved when the basis vectors $\mathbf{v}_i$ are acted on by a unitary transformation matrix. 



The purpose of the Jacobi method is to find the eigenvalues and the corresponding eigenvectors of the system. When the unitary transformation is performed, the eigenvalues and the eigenvectors of the matrix equation is conserved. The goal of the Jacobi method is to transform an arbitrary dense matrix to a diagonal matrix. Then, the diagonal elements will be the eigenvalues of the matrix equation. The same transformations are performed on the right hand side of the equations, and here we are left with the eigenvectors of the matrix.

 A more detailed explanation of the Jacobi method  is given in the lecture notes of this course\footnote{Morten Hjorth-Jensen \textit{Computational Physics: Lecture Notes Fall 2015 }, 215-218} 
\subsection{Numerical implementation}


The following block of code applies Jacobi's method as long as the number of iterations is smaller than the maximum number of iterations and the maximum non-diagonal element is bigger than the tolerance, $\epsilon$. The maximum of iterations is put to $N^3$, where $N$ is the dimension of the matrices and $\epsilon = 10^{-8}$. 

%oda er dust og forstår ikke mappestrukturen sin, unnskyld git
\begin{lstlisting}
 while (num_iter <= max_iter && fabs(max) > epsilon)
        {
            max = max_elem(&i, &j, A, N);
            A = sim_transform(i, j, A, R,N);
            num_iter++;
        }
\end{lstlisting}
The \lstinline{max_elem}-function in the code above finds the maximum non-diagonal element of the matrix. As long as the absolute value of this element is not smaller than the tolerance, $\epsilon$, and the max number of iterations has not been reached, it will continue to loop, and calculate a new matrix using the \lstinline{sim_transform}-function. It is in this function that the Jacobi transformation is performed. The method was implemented as follows:
\newline
\begin{lstlisting}
   if (A(k, l) != 0.0){
        tau = (A(l,l) - A(k,k))/(2.0*A(k,l));

        if(tau > 0){
            t = 1.0/(tau + sqrt(1.0 + tau*tau));
            //t = +tau - sqrt(1+tau*tau);
        }
        else{
            t = -1.0/(-tau + sqrt(1.0 + tau*tau));
        }
        c = 1/(sqrt(1.0 +t*t));
        s = c*t;
    }
    else{
        c = 1.0;
        s = 0.0;
    }
    double a_kk, a_ll, a_ik, a_il, r_ik, r_il;
    a_kk = A(k,k);
    a_ll = A(l,l);
    A(k,k) = c*c*a_kk - 2.0*c*s*A(k,l) + s*s*a_ll;
    A(l,l) = s*s*a_kk + 2.0*c*s*A(k,l) + c*c*a_ll;
    A(k,l) = 0.0;  // hard-coding non-diagonal elements by hand
    A(l,k) = 0.0;  // same here
    for ( int i = 0; i < N+1; i++ ) {
      if ( i != k && i != l ) {
        a_ik = A(i,k);
        a_il = A(i,l);
        A(i,k) = c*a_ik - s*a_il;
        A(k,i) = A(i,k);
        A(i,l) = c*a_il + s*a_ik;
        A(l,i) = A(i,l);
      }
      r_ik = R(i,k);
      r_il = R(i,l);

      R(i,k) = c*r_ik - s*r_il;
      R(i,l) = c*r_il + s*r_ik;
\end{lstlisting}

\end{document}
